### 2018-10-16 Manuscript/Figures/LIP_Table.tex
calculate individual halflives BEFORE rounding, and update figure labels

\begin{tabular}{lrlrrlrrl}
\toprule
                name &    age &                     age ref. &  original area Mkm2 &  present area Mkm2 &                                 present area ref. &  present/original &  half-life &   buried \\
\midrule
      Columbia River &   16.0 &   Kasbohm and Schoene (2018) &                0.68 &               0.38 &                           Buchan and Ernst (2004) &              0.56 &       19.2 &       no \\
                Afar &   30.0 &  Courtillot and Renne (2003) &                2.05 &               0.63 &                              Coffin et al. (2006) &              0.31 &       17.7 &  partial \\
                NAIP &   62.0 &         Larsen et al. (2016) &                1.07 &               0.29 &     Buchan and Ernst (2004), Coffin et al. (2006) &              0.27 &       33.0 &  partial \\
              Deccan &   66.0 &               Schoene (2015) &                0.83 &               0.56 &                              Coffin et al. (2006) &              0.68 &      116.6 &       no \\
          Seychelles &   66.0 &               Schoene (2015) &                0.46 &               0.00 &                              Coffin et al. (2006) &              0.00 &        0.0 &      yes \\
          Madagascar &   90.0 &    Cucciniello et al. (2010) &                0.63 &               0.03 &                              Coffin et al. (2006) &              0.05 &       20.8 &       no \\
 Caribbean-Colombian &   94.0 &         Loewen et al. (2013) &                0.71 &               0.13 &                              Coffin et al. (2006) &              0.18 &       37.6 &       no \\
               HALIP &   95.0 &      Kingsbury et al. (2018) &                3.60 &               0.15 &                      Moosdorf and Hartmann (2015) &              0.04 &       20.7 &       no \\
              EQUAMP &  131.0 &       Hollanda et al. (2016) &                0.66 &               0.01 &                            Hollanda et al. (2016) &              0.01 &       20.5 &       no \\
               Comei &  132.0 &            Zhu et al. (2009) &                0.11 &                NaN &                                               n/a &               NaN &        NaN &       no \\
             Bunbury &  132.0 &            Zhu et al. (2009) &                0.03 &               0.00 &                              Thorne et al. (2014) &              0.05 &       30.9 &       no \\
     Parana-Etendeka &  135.0 &      Florisbal et al. (2014) &                3.12 &               0.40 &  Coffin et al. (2016), Teixeira and Albrez (2017) &              0.13 &       45.7 &  partial \\
                Trap &  140.0 &          Ernst et al. (2001) &                0.03 &               0.00 &                           Ernst and Buchan (2001) &              0.00 &        0.0 &       no \\
 NW Australia Margin &  160.0 &            COME BACK TO THIS &                0.62 &               0.00 &                              Coffin et al. (2006) &              0.00 &        0.0 &      yes \\
               Karoo &  183.0 &        Burgess et al. (2015) &                3.21 &               0.15 &                               de Kock compilation &              0.05 &       41.3 &       no \\
              Ferrar &  183.0 &        Burgess et al. (2015) &                0.18 &                NaN &                                               n/a &               NaN &        NaN &       no \\
                CAMP &  201.0 &      Blackburn et al. (2013) &               11.46 &               0.23 &                   Marzoli and Parisio compilation &              0.02 &       35.7 &  partial \\
             Siberia &  252.0 &   Burgess and Bowring (2015) &                3.46 &               0.47 &                              Coffin et al. (2006) &              0.14 &       87.5 &       no \\
            Emeishan &  259.0 &           Zhou et al. (2002) &                0.71 &               0.06 &                              Coffin et al. (2006) &              0.09 &       72.9 &       no \\
    Panjal-Qiangtang &  283.0 &           Zhai et al. (2013) &                0.11 &                NaN &                                               n/a &               NaN &        NaN &       no \\
               Tarim &  290.0 &             Xu et al. (2014) &                0.35 &                NaN &                                               n/a &               NaN &        NaN &       no \\
            Magdalen &  360.0 &         Murphy et al. (1999) &                0.42 &                NaN &                                               n/a &               NaN &        NaN &       no \\
              Vilyui &  374.0 &          Renne et al. (2010) &                1.14 &                NaN &                                               n/a &               NaN &        NaN &       no \\
        Kola-Dnieper &  380.0 &    Arzamastsev and Wu (2014) &                5.90 &                NaN &                                               n/a &               NaN &        NaN &       no \\
            Suordakh &  450.0 &       Khudoley et al. (2013) &                0.02 &                NaN &                                               n/a &               NaN &        NaN &       no \\
         Kalkarindji &  511.0 &        Jourdan et al. (2014) &                3.54 &               0.17 &                              Thorne et al. (2014) &              0.05 &      116.3 &       no \\
            Franklin &  720.0 &       Denyszyn et al. (2009) &                2.62 &               0.04 &                           Buchan and Ernst (2004) &              0.02 &      121.8 &       no \\
\bottomrule
\end{tabular}